Diagnosing Simple Wind Models of Cool Gas Outflows through Metal-Line Absorption and Emission

### 2010-10-06 Winds/papers/Theory/First/Submit/ms.tex
profx
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5257 ac11df5e-a788-c23b-a8e1-c758c323924d

\newcommand{\nhn}{$n_{\rm H}^0$}
\newcommand{\mnhn}{n_{\rm H}^0}
\newcommand{\mnhf}{n_{\rm H}^{0,f}}
\newcommand{\msun}{M_\odot}
\newcommand{\tpk}{$\tau_{\rm pk}$}
\newcommand{\mtpk}{\tau_{\rm pk}}
\newcommand{\mdv}{\delta v}
\newcommand{\ewabs}{$W_{\rm a}$}
\newcommand{\ewe}{$W_{\rm e}$}
\newcommand{\rtt}{$r_{\tau = 0.2}$}
\newcommand{\mrtt}{r_{\tau = 0.2}}
\newcommand{\bturb}{$b_{\rm turb}$}
\newcommand{\mbturb}{b_{\rm turb}}
\newcommand{\taud}{$\tau_{\rm dust}$}
\newcommand{\mtaud}{\tau_{\rm dust}}
\newcommand{\maconfig}{2p$^6$3s}
\newcommand{\mbconfig}{2p$^6$3p}
\newcommand{\aconfig}{a~$^6$D$^0$}
\newcommand{\zconfig}{z~$^6$D$^0$}
\newcommand{\mvr}{v_{\rm r}}
\newcommand{\naid}{\ion{Na}{1}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\newcommand{\feiia}{\ion{Fe}{2}~$\lambda 2586$}
\newcommand{\feiib}{\ion{Fe}{2}~$\lambda 2600$}
\newcommand{\feiic}{\ion{Fe}{2}*$ \; \lambda 2612$}
\newcommand{\feiie}{\ion{Fe}{2}*$ \; \lambda 2626$}
\newcommand{\feiid}{\ion{Fe}{2}~$\lambda\lambda 2586, 2600$}
\newcommand{\feiis}{\ion{Fe}{2}*}
\newcommand{\nmg}{$n_{\rm Mg^+}$}
\newcommand{\mnmg}{n_{\rm Mg^+}}
\newcommand{\nfe}{$n_{\rm Fe^+}$}
\newcommand{\mnfe}{n_{\rm Fe^+}}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {\, {\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

\title{Idealized Wind Models of Cool Gas Outflows}

@@

\begin{abstract}
\begin{itemize}
\item Emission is a generic feature of (nearly) isotropic winds.  This
  emission fills in absorption at $v \approx 0$, significantly
  complicating the absorption-line analysis, especially of an ISM
  component. 
\item The relative strengths of the emission
lines, therefore, is sensitive to both the opacity and velocity
extent of the wind.
\item  One must take careful
consideration of line emission, therefore,  before drawing robust
conclusions about the presence/absence of gas with small velocity
offsets from systemic. 
\end{itemize}
\end{abstract}

@@

The observation of metal-line absorption is now a well-established
means of identifying outflows.  Furthermore, because the X-ray and
H$\alpha$ emission generated by winds is faint, absorption-line
analyses have traditionally been the only way to probe outflows in
distant galaxies. However, very little research
has been directed
toward comparing the observations against (even idealized) wind
models (although see \citealt{fmm+09}).  
Instead, researchers have gleaned what limited information
is afforded by direct analysis of the absorption lines.  The data
diagnose the speed of the gas relative to the galaxy, 
yet they poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the flow from the galaxy.   In turn, constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of
galactic-scale winds, therefore, remain open matters of debate
\citep{mqt05,sdr08,sh09,ssr02,od06,kkd+09}.

Recent studies of $z \lesssim 1$ star-forming galaxies have revealed that
the cool outflowing gas often exhibits significant resonant-line emission (e.g.\
\ion{Mg}{2}, \ion{Na}{1}) in
tandem with the nearly ubiquitous blue-shifted absorption
\citep{wcp+09,rwk+10,cth+10}.  The resultant spectra resemble the P-Cygni
profile characteristic of stellar winds.
This phenomenon was first reported by \citet{phillips93}, who observed 
blue-shifted absorption and red-shifted emission in the \ion{Na}{1}
transition in the spectrum of the local starburst galaxy NGC 1808.
More recently, \cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emission, both directly and when the authors modeled and
`removed' the $v \approx 0\mkms$ absorption component.  Although the
authors suggested the emission could be related to back-scattered
light in the wind, they also allowed that it could be related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rwk+10} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies.

Bright \ion{Mg}{2} line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite[][see also
Rubin et al\ 2011, in prep.]{rubin+10c}.
In their analysis of a single galaxy spectrum, \cite{rubin+10c}
further demonstrated that the \ion{Mg}{2} line emission is spatially
extended, and used the size of the emission to infer that the wind 
extends to at least 7\,kpc from the galaxy.
These authors additionally detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions, and attribute
the emission to fluorescence powered by \ion{Fe}{2} resonant
absorption.  In other words, these photons are re-emitted by the wind
into our sightline, and are analogous to the emitted photons in a P-Cygni profile.
%Their interpretation is that the emission is the result of photons
%scattered off the wind into our sightline, analogous to P-Cygni.
%\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
%which they use to estimate the size of the outflow.  They infer that
%the wind extends to at least 7\,kpc from the galaxy.  
Line-emission is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions \citep{prs+02,shapley03}.
We emphasize that this emission arises from a completely different
physical process than those generating X-ray and H$\alpha$ emission
(e.g., shocks), and presumably probes both the cool gas at the base of
the wind and the outskirts of the flow (i.e., wherever a given
transition is optically thick).
A comprehensive analysis of the %emission lines
scattered and fluorescent emission
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit [IFU] observations,
\citealt{ssb+05,wsg08}) 
may offer unique
diagnostics on the spatial extent, morphology, and density of the outflow 
from distant galaxies,
eventually
setting tighter constraints on the energetics of the flow.

@@

Although astronomers are rapidly producing a wealth of observational
datasets on galactic-scale winds, a key ingredient to a proper
analysis is absent.  Just as comparisons between observed supernova 
lightcurves and spectra and radiative transfer calculations of
of SNe models have provided crucial insight into the physics driving,
e.g., standard candle relationships for both Type Ia and IIP SNe 
\citep{kw07,kw09}, modeling of the observable signatures of galactic
outflows is necessary for understanding the physical
properties of the gas and ultimately the physics driving the flows. 
%[Draw analogy to exploring SN lightcurves and spectra]
In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows as observed in one
dimensional spectra and by IFUs.  Using Monte Carlo
radiative transfer techniques, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties, accounting for the effects of resonant scattering and
fluorescence. 
Although the winds are idealized, the results frequently
contradict our intuition and 
challenge the straightforward conversion of observables to (even crude) physical
constraints.  We caution that these findings have a direct bearing on
recent and upcoming surveys of galactic outflows, particularly those
which make use of \ion{Mg}{2} and \ion{Fe}{2} transitions to probe
outflow properties.
%[Add another line or two]

@@

[JXP:  This needs to be re-run and revised following Kate's update.
Will do so in the next few days (hopefully tomorrow)]

@@

Iron exhibits the most complex set of energy levels for elements
frequently studied in astrophysics.  The Fe$^+$ ion alone has 
millions of energy levels recorded \citep{kurucz_cd1}, and even this is an
incomplete list.  
One reason for iron's complexity is
that the majority of its configurations exhibit fine-structure splitting.
This includes the ground-state configuration (\aconfig) which is split
into 5 levels, 
labeled by the total angular momentum $J$, 
with excitation energies $T_{\rm ex} \equiv \Delta E / k$ ranging from
$T_{\rm ex} \approx 500-1500$\,K (Figure~\ref{fig:energy}).  
Transitions between these fine-structure levels are 
forbidden (magnetic-dipole) and have spontaneous decay times of several hours.

@@

Nearly all of the absorption studies of galactic-scale outflows 
have focused on intensely, star-forming galaxies.  The
intrinsic emission of these galaxies is a complex combination of
stars and \ion{H}{2} regions that is then modulated by dust and gas
within the interstellar medium (ISM).  For the spectral regions studied
here, most stars show a featureless continuum, but a few spectral
types do show significant \ion{Mg}{2} and \ion{Fe}{2} absorption.
In addition, asymmetric and/or blue-shifted absorption is exhibited in these
transitions in A and F stars driving stellar winds \citep{sll+94}.
\ion{Mg}{2} P Cygni profiles are observed in a handful of F stars by \cite{sll+94}, who
attribute the emission to chromospheric activity rather than mass-loss effects.
%{\bf [Kate: Any emission??]}  
\ion{H}{2} regions, meanwhile, are observed to emit weakly
at the \mgiid\ doublet, primarily due to recombinations in the outer
layers \citep{kbc+93} {\bf [Kate: How about FeII?  X: I don't know of any examples of this.]}.  
It is beyond the scope of this paper to
properly model the 
stellar absorption and \ion{H}{2} region emission, but the reader
should be aware that they can complicate the observed spectrum,
independent of the outflow, especially at velocities $v \approx 0 \mkms$.
In the following, we assume a simple flat continuum 
normalized to unit value.  The size of the emitting
region $r_{\rm source}$ is a free parameter, but we restrict its value
to be smaller than the minimum radial extent of any gaseous
component.   Lastly, the source does not absorb any scattered or
emitted photons.

@@

We calculated spectra using a 3D Monte Carlo radiation transport code originally designed for supernova outflows  \cite{Kasen_2006} but  modified to treat resonant line transport  on a galactic scale \cite{Kasen_lyman}.  Our methods are similar to those used in several other codes \citep[e.g.,][]{Zheng_2002,Dijkstra_2006,Verhame_2006,Laursen_2009}, except that here we include the effects of multiple line scattering and fluorescence.

@@

Photon packets were tracked through randomized scattering and absorption events until they escaped the computational domain.  The distance to the next packet interaction event was determined by Doppler shifting packets to the comoving frame and sampling the mean free path to each resonance line, as well to dust.  The resonant line opacity followed a Voigt profile which was determined using the analytic fits of  \cite{Tomi_2006}.      Non-resonant lines were assumed to be completely optically thin, and the dust opacity was assumed to be wavelength independent and completely absorbing (right X?).   Because of the relatively low optical depths encountered in these models, no ``core-skipping'' scheme was applied to accelerate the resonant line transport,  and recoil effects were ignored.

@@

To validate the monte carlo code, we calculated a series of test problems consisting of a point source embedded in a spherical, homogenous medium of varying resonant line optical depth.  The resulting resonant line profiles  \citep[see][]{kasen_lyman} were
found to be in good agreement with the standard analytic solutions \citep{Harrington_1973,Neufeld_1990,Dijkstra_2006}.  As a further test, we calculated line profiles for the case of an extended spherical source embedded in a homologously expanding wind with large velocity gradient.  These model spectra (Figure~\ref{fig:oneline_test}) are in good agreement with ones determined by direct numerical integration of the radiation transport equation under the Sobolev approximation [e.g.,][]\citep{Jeffery_Branch}.

@@

For the majority of models studied in this paper, we assume the gas
contains no dust.
This is an invalid assumption, especially for material associated with
the ISM of a galaxy.  
Extraplanar material likely associated with a galactic-scale 
outflow has been observed to emit IR radiation characteristic of dust 
\citep[e.g.][]{M87_dust,hrg90,adb99,rkl01}.
Essentially all astrophysical environments that contain both cool gas
and metals also show signatures of dust depletion and extinction.  This includes the
ISM of star-forming and \ion{H}{1}-selected galaxies
\citep[e.g.][]{ss96,pw01,pcd+07}, strong \ion{Mg}{2} metal-line
absorption systems \citep{ykv+06,mnt+08}, and the galactic winds traced
by low-ion transitions \citep{prs+02,rvs05b}.  Although the galactic winds
traced specifically by \ion{Mg}{2} and \ion{Fe}{2} transitions have not (yet) been
demonstrated to contain dust, it is reasonable to consider its
effects.
In addition, \cite{msf+10} have argued from a statistical analysis
that dust is distributed to many tens kpc from $z \sim 0.1$ galaxies
and have suggested it was transported from the galaxies by
galactic-scale winds.  One is motivated, therefore, to consider
models that include the effects of dust extinction.

For analysis on normalized spectra, the effects of dust are largely minimized; 
dust has a nearly constant opacity over small spectral regions and all
features are simply scaled together.
For scattered and resonantly trapped photons, however, the relative effect of dust
extinction can be much greater.  
These photons travel a much longer distance to escape the medium and
may experience a much higher integrated opacity from dust.
Indeed, dust is frequently invoked to
explain the weak (or absent) \lya\ emission from star-forming galaxies
\citep[e.g.][]{shapley03}.  Although the transitions studied here have
much lower opacity than \lya, dust could still play an important role
in the predicted profiles.

In a few models, we include absorption by dust under the following
assumptions:
(i) the dust opacity scales with the density of the gas (i.e.\ we
adopt a fixed dust-to-gas ratio);
(ii) the opacity is independent of wavelength, a reasonable
approximation given the small spectral range analyzed;
(iii) the photons absorbed by dust are re-emitted at IR wavelengths
and are `lost' from the system.  The dust absorption is normalized by \taud, 
the integrated opacity of dust from the center of the system to
infinity.  The ambient ISM of a star-forming galaxy may be expected to exhibit
\taud\ values of one to a few at $\lambda \sim 3000$\AA\ \citep[e.g.][]{dust_again,cf00}.

@@

Similar results are observed for the \ion{Fe}{2} resonance
transitions (Figure~\ref{fig:fiducial_ifu_feii}).
For transitions to fine-structure levels of the \aconfig, the source
is unattenuated but there is a significant contribution from photons
generated in the wind. 
The results presented in Figures~\ref{fig:fiducial_ifu_mgii} and
\ref{fig:fiducial_ifu_feii} are sensitive to the radial extent,
morphology, density and velocity profiles of this galactic-scale
wind.  Consequently, IFU observations of line emission from low-ion
transitions may offer the most direct constraints on galactic-scale
wind properties.

>>>>

\newcommand{\nhn}{$n_{\rm H}^0$}
\newcommand{\mnhn}{n_{\rm H}^0}
\newcommand{\mnhf}{n_{\rm H}^{0,f}}
\newcommand{\msun}{M_\odot}
\newcommand{\tpk}{$\tau_{\rm pk}$}
\newcommand{\mtpk}{\tau_{\rm pk}}
\newcommand{\mdv}{\delta v}
\newcommand{\ewabs}{$W_{\rm a}$}
\newcommand{\ewe}{$W_{\rm e}$}
\newcommand{\rtt}{$r_{\tau_S = 0.2}$}
\newcommand{\mrtt}{r_{\tau_S = 0.2}}
\newcommand{\bturb}{$b_{\rm turb}$}
\newcommand{\mbturb}{b_{\rm turb}}
\newcommand{\taud}{$\tau_{\rm dust}$}
\newcommand{\mtaud}{\tau_{\rm dust}}
\newcommand{\maconfig}{2p$^6$3s}
\newcommand{\mbconfig}{2p$^6$3p}
\newcommand{\aconfig}{a~$^6$D$^0$}
\newcommand{\zconfig}{z~$^6$D$^0$}
\newcommand{\mvr}{v_{\rm r}}
\newcommand{\naid}{\ion{Na}{1}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\newcommand{\feiia}{\ion{Fe}{2}~$\lambda 2586$}
\newcommand{\feiib}{\ion{Fe}{2}~$\lambda 2600$}
\newcommand{\feiic}{\ion{Fe}{2}*$ \; \lambda 2612$}
\newcommand{\feiie}{\ion{Fe}{2}*$ \; \lambda 2626$}
\newcommand{\feiid}{\ion{Fe}{2}~$\lambda\lambda 2586, 2600$}
\newcommand{\feiis}{\ion{Fe}{2}*}
\newcommand{\nmg}{$n_{\rm Mg^+}$}
\newcommand{\mnmg}{n_{\rm Mg^+}}
\newcommand{\nfe}{$n_{\rm Fe^+}$}
\newcommand{\mnfe}{n_{\rm Fe^+}}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {\, {\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

\title{Diagnosing Simple Wind Models of Cool Gas Outflows through
  Metal-Line Absorption and Emission}

@@

\begin{abstract}
We analyze the absorption and emission-line profiles produced by a set
of simple, cool gas wind models motivated by galactic-scale outflow
observations.  We focus on the \mgiid\ doublet and \ion{Fe}{2} UV1
multiplet at $\lambda \approx 2600$\AA, but the results are applicable
to most other transitions used to study outflows (e.g.\ \ion{Na}{1},
\ion{H}{1} \lya, \ion{Si}{2}).   We implement monte carlo radiative
transfer techniques to properly track the propogation of scattered and
flourescent photons to [produce] 1D spectra and 2D spectral images.
[Should we make the codes publically available?  Upon request?]
[Mention fiducial model here?]
By design, the \ion{Mg}{2} doublet exhibits strong, blue-shifted
absorption but also comparably strong line-emission at roughly
systemic velocity.  This line-emission `fills-in' the absorption at
velocieties near systemic, significantly reducing the inferred
opacity. Analysis of cool gas outflows that entirely ignores this
line-emission may incorrectly conclude that the gas is partially
covered, has a significantly lower peak optical depth, and/or that a
gas with velocities near systemic (e.g.\ from the ISM or even
an infalling component) is absent.   
The \ion{Fe}{2} resonance lines show similar behaviour, however 
these lines are connected to fine-structure levels that are optically
thin to the medium.  As a result, the resonant line-emission is
reduced (at the `expense' of significant fine-structure emission) and
the absorption lines more faithfully reproduce the intrinsic opacity
of the wind.   [Together these results naturally explain the
absorption and emission-line characteristics observed for star-forming
galaies at $z<1$.]  
We examine outflows with a range of power-law density and velocity
profiles and a radiative-driven wind model.  The resultant \ion{Mg}{2}
and \ion{Fe}{2} profiles show a diverse complexity of absorption and
emission characteristics but overall have qualitatively similar
features.  We also study a scenario purported to describe the outflows
of $z \sim 3$ Lyman break galaxies and argue this model fails to
reproduce the observations when (previously ignored0 scattered photons
are properly considered.   Strong line-emission from metal-line
transitions is a generic prediction of wind models, even in the
presence of dust and anisotropic configurations.  We calculate the
surface brightness profiles of this emission which motivates
spatially-extended spectroscopic observations with current and
next-generation integral-field-unit instruments.

\end{abstract}

@@

The observation of metal-line absorption is now a well-established
means of identifying outflows.  Furthermore, because the X-ray and
H$\alpha$ emission generated by winds is faint, absorption-line
analyses have traditionally been the only way to probe outflows in
distant galaxies. However, very little research
has been directed
toward comparing the observations against (even idealized) wind
models \citep[e.g.][]{fmm+09}.  
Instead, researchers have gleaned what limited information
is afforded by direct analysis of the absorption lines.  The data
diagnose the speed of the gas relative to the galaxy, 
yet they poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the flow from the galaxy.   In turn, constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of
galactic-scale winds, therefore, remain open matters of debate
\citep{mqt05,sdr08,sh09,ssr02,od06,kkd+09}.

Recent studies of $z \lesssim 1$ star-forming galaxies have revealed that
the cool outflowing gas often exhibits significant resonant-line emission (e.g.\
\ion{Mg}{2}, \ion{Na}{1}) in
tandem with the nearly ubiquitous blue-shifted absorption
\citep{wcp+09,rwk+10,cth+10}.  The resultant spectra resemble the P-Cygni
profile characteristic of stellar winds.
This phenomenon was first reported by \citet{phillips93}, who observed 
blue-shifted absorption and red-shifted emission for the \ion{Na}{1}
transition in the spectrum of the local starburst galaxy NGC 1808.
More recently, \cite{wcp+09}, who
studied \ion{Mg}{2} absorption in $z \sim 1.4$ galaxies, reported
\ion{Mg}{2} emission in a small subset of the individual galaxy
spectra of their large sample.  These were excluded from the full
analysis on concerns that the emission was related to AGN activity.
The stacked spectra of the remaining galaxies, however, also indicated
\ion{Mg}{2} emission, both directly and when the authors modeled and
`removed' the $v \approx 0\mkms$ absorption component.  The
authors suggested the emission could be related to back-scattered
light in the wind, but presumed that it was related to weak
AGN activity.   Very similar \ion{Mg}{2} emission was observed by
\cite{rwk+10} who repeated the analysis of \cite{wcp+09} on a set
of lower redshift galaxies.

Bright \ion{Mg}{2} line emission has also been
reported for individual galaxies at $z \sim 0.7$ by \cite[][see also
Rubin et al\ 2011, in prep.]{rubin+10c}.
In their analysis of a single galaxy spectrum, \cite{rubin+10c}
further demonstrated that the \ion{Mg}{2} line emission is spatially
extended, and used the size of the emission to infer that the wind 
extends to at least 7\,kpc from the galaxy.
These authors additionally detected line
emission from non-resonant \ion{Fe}{2}$^*$ transitions, and attribute
the emission to fluorescence powered by \ion{Fe}{2} resonant
absorption.  In other words, these photons are re-emitted by the wind
into our sightline, and are analogous to the emitted photons in a P-Cygni profile.
%Their interpretation is that the emission is the result of photons
%scattered off the wind into our sightline, analogous to P-Cygni.
%\cite{rubin+10c} also observed spatially extended \ion{Mg}{2} emission
%which they use to estimate the size of the outflow.  They infer that
%the wind extends to at least 7\,kpc from the galaxy.  
Line-emission is
also observed for $z \sim 3$ LBGs in the resonant \lya\ transition
and non-resonant \ion{Si}{2}$^*$ transitions\footnote{See also
  \cite{france10} for a new example at $z \sim 0$.} \citep{prs+02,shapley03}.
We emphasize that this emission arises from a completely different
physical process than those generating X-ray and H$\alpha$ emission
(e.g., shocks), and presumably probes both the cool gas at the base of
the wind and the outskirts of the flow (i.e., wherever a given
transition is optically thick).
A comprehensive analysis of the %emission lines
scattered and fluorescent emission
related to galactic-scale outflows
(e.g.\ via deep integral-field-unit [IFU] observations,
\citealt{ssb+05,wsg08}) 
may offer unique
diagnostics on the spatial extent, morphology, and density of the outflow 
from distant galaxies,
eventually
setting tighter constraints on the energetics of the flow.

@@

Although astronomers are rapidly producing a wealth of observational
datasets on galactic-scale winds, a key ingredient to a proper
analysis is absent.  Just as comparisons between observed supernova 
lightcurves and spectra and radiative transfer calculations of
of SNe models have provided crucial insight into the physics driving,
e.g., standard candle relationships for both Type Ia and IIP SNe 
\citep{kw07,kw09}, modeling of the observable signatures of galactic
outflows is necessary for understanding the physical
properties of the gas and ultimately the physics driving the flows. 
%[Draw analogy to exploring SN lightcurves and spectra]
In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows as observed in one
dimensional spectra and by IFUs.  Using Monte Carlo
radiative transfer techniques, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties, accounting for the effects of resonant scattering and
fluorescence. 
Although the winds are idealized, the results frequently
contradict our intuition and 
challenge the straightforward conversion of observables to (even crude) physical
constraints.  These findings have a direct bearing on
recent and upcoming surveys of galactic outflows, particularly those
which make use of \ion{Mg}{2} and \ion{Fe}{2} transitions to probe
outflow properties.
%[Add another line or two]

@@

@@

Iron exhibits the most complex set of energy levels for elements
frequently studied in astrophysics.  The Fe$^+$ ion alone has 
millions of energy levels recorded \citep{kurucz05}, and even this is an
incomplete list.  
One reason for iron's complexity is
that the majority of its configurations exhibit fine-structure splitting.
This includes the ground-state configuration (\aconfig) which is split
into 5 levels, 
labeled by the total angular momentum $J$, 
with excitation energies $T_{\rm ex} \equiv \Delta E / k$ ranging from
$T_{\rm ex} \approx 500-1500$\,K (Figure~\ref{fig:energy}).  
Transitions between these fine-structure levels are 
forbidden (magnetic-dipole) and have spontaneous decay times of several hours.

@@

Nearly all of the absorption studies of galactic-scale outflows 
have focused on intensely, star-forming galaxies.  The
intrinsic emission of these galaxies is a complex combination of
stars and \ion{H}{2} regions that is then modulated by dust and gas
within the interstellar medium (ISM).  For the spectral regions studied
here, most stars show a featureless continuum, but a few spectral
types do show significant \ion{Mg}{2} and \ion{Fe}{2} absorption.
In addition, asymmetric and/or blue-shifted absorption is exhibited in these
transitions in A and F stars driving stellar winds \citep{sll+94}.
\ion{Mg}{2} P Cygni profiles are observed in a handful of F stars by \cite{sll+94}, who
attribute the emission to chromospheric activity rather than mass-loss effects.
\ion{H}{2} regions, meanwhile, are observed to emit weakly
at the \mgiid\ doublet, primarily due to recombinations in the outer
layers \citep{kbc+93}.
It is beyond the scope of this paper to
properly model the 
stellar absorption and \ion{H}{2} region emission, but the reader
should be aware that they can complicate the observed spectrum,
independently of any outflow, especially at velocities $v \approx 0 \mkms$.
In the following, we assume a simple flat continuum 
normalized to unit value.  The size of the emitting
region $r_{\rm source}$ is a free parameter, but we restrict its value
to be smaller than the minimum radial extent of any gaseous
component.   Lastly, the source does not absorb any scattered or
emitted photons.

@@

We calculated spectra using a 3D Monte Carlo radiation transport code
originally designed for supernova outflows  \citep{Kasen_2006} but
modified to treat resonant line transport  on a galactic scale
\citep{Kasen_lyman}.  Our methods are similar to those used in several
other codes which focus on \lya\
\citep[e.g.,][]{Zheng_2002,Dijkstra_2006,Verhame_2006,Laursen_2009},
except that here we include the effects of multiple line scattering
and fluorescence.

@@

Photon packets were tracked through randomized scattering and
absorption events until they escaped the computational domain.  The
distance to the next packet interaction event was determined by
Doppler shifting packets to the comoving frame and sampling the mean
free path to each resonance line (and dust, if present).  The resonant
line opacity followed a Voigt profile which was determined using the
analytic fits of  \cite{Tomi_2006}.      Non-resonant lines were
assumed to be completely optically thin, and the dust opacity was
assumed to be wavelength independent and completely absorbing
($\S$~\ref{sec:dust_method}).
Because of the relatively low optical depths encountered in
these models, no ``core-skipping'' scheme was applied to accelerate
the resonant line transport,  and recoil effects were ignored.

@@

To validate the monte carlo code, we calculated a series of test problems consisting of a point source embedded in a spherical, homogenous medium of varying resonant line optical depth.  The resulting resonant line profiles  \citep[see][]{Kasen_lyman} were
found to be in good agreement with the standard analytic solutions
\citep{Harrington_1973,Neufeld_1990,Dijkstra_2006}.  As a further
test, we calculated line profiles for the case of an extended
spherical source embedded in a homologously expanding wind with large
velocity gradient.  These model spectra
(Figure~\ref{fig:oneline_test}) are in good agreement with ones
determined by direct numerical integration of the radiation transport
equation under the Sobolev approximation \citep[e.g.,][]{Jeffery_Branch}.

@@

For the majority of models studied in this paper, we assume the gas
contains no dust.
This is an invalid assumption, especially for material associated with
the ISM of a galaxy.  
Essentially all astrophysical environments that contain both cool gas
and metals also show signatures of dust depletion and extinction.  This includes the
ISM of star-forming and \ion{H}{1}-selected galaxies
\citep[e.g.][]{ss96,pw01,pcd+07}, strong \ion{Mg}{2} metal-line
absorption systems \citep{ykv+06,mnt+08}, and the galactic winds traced
by low-ion transitions \citep{prs+02,rvs05b}.  
Extraplanar material likely associated with a galactic-scale 
outflow has been observed to emit IR radiation characteristic of dust 
\citep[e.g.][]{hrg90,adb99,rkl01}.
In addition, \cite{msf+10} have argued from a statistical analysis
that dust is distributed to many tens kpc from $z \sim 0.1$ galaxies
and have suggested it was transported from the galaxies by
galactic-scale winds.  
Although the galactic winds
traced specifically by \ion{Mg}{2} and \ion{Fe}{2} transitions have not (yet) been
demonstrated to contain dust, it is reasonable to consider its effects.

For analysis on normalized spectra (i.e.\ absorption lines), the effects of dust are largely minimized; 
dust has a nearly constant opacity over small spectral regions and all
features are simply scaled together.
For scattered and resonantly trapped photons, however, the relative effect of dust
extinction can be much greater.  
These photons travel a much longer distance to escape the medium and
may experience a much higher integrated opacity from dust.
Indeed, dust is frequently invoked to
explain the weak (or absent) \lya\ emission from star-forming galaxies
\citep[e.g.][]{shapley03}.  Although the transitions studied here have
much lower opacity than \lya, dust could still play an important role
in the predicted profiles.

In a few models, we include absorption by dust under the following
assumptions:
(i) the dust opacity scales with the density of the gas (i.e.\ we
adopt a fixed dust-to-gas ratio);
(ii) the opacity is independent of wavelength, a reasonable
approximation given the small spectral range analyzed;
(iii) dust absorbs but does not scatter photons;
(iv) the photons absorbed by dust are re-emitted at IR wavelengths
and are `lost' from the system.  The dust absorption is normalized by \taud, 
the integrated opacity of dust from the center of the system to
infinity.  The ambient ISM of a star-forming galaxy may be expected to exhibit
\taud\ values of one to a few at $\lambda \sim 3000$\AA\ \citep[e.g.][]{cf00}.

@@

Similar results are observed for the \ion{Fe}{2} resonance
transitions (Figure~\ref{fig:fiducial_ifu_feii}).
For transitions to fine-structure levels of the \aconfig\
configuration, the source
is unattenuated but there is a significant contribution from photons
generated in the wind. 
The results presented in Figures~\ref{fig:fiducial_ifu_mgii} and
\ref{fig:fiducial_ifu_feii} are sensitive to the radial extent,
morphology, density and velocity profiles of this galactic-scale
wind.  Consequently, IFU observations of line emission from low-ion
transitions may offer the most direct constraints on galactic-scale
wind properties.
